---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: console
---
Data Analysis through visualisation in R

First, importing the necessary libraries in R. 

ggplot: Creating aesthetic visualisation plots
lubridate: Dealing with time-frames in the dataset
dplyr: Data Manipulation
tidyr: Tidy Messy Data 
DT: Datatables in JavaScript
scales: Map data to correct scales with axes and legends.
```{r}
library(ggplot2)
library(ggthemes)
library(lubridate)
library(dplyr)
library(tidyr)
library(DT)
library(scales)
```

Reading the data from the csv files and then combining together
```{r}

april <- read.csv('uber-raw-data-apr14.csv')
may <- read.csv('uber-raw-data-may14.csv')
june <- read.csv('uber-raw-data-jun14.csv')
july <- read.csv('uber-raw-data-jul14.csv')
aug <- read.csv('uber-raw-data-aug14.csv')
sept <- read.csv('uber-raw-data-sep14.csv')
data <- rbind(april, may, june, july, aug, sept)

```

Here, we will convert the Date.Time into a more readable format using Date time conversion
```{r}

data$Date.Time <- as.POSIXct(data$Date.Time, format= "%m/%d/%Y %H:%M:%S")
data$Time <- format(as.POSIXct(data$Date.Time, format = "%m/%d/%Y %H:%M:%S"), format="%H:%M:%S")
data$Date.Time <- ymd_hms(data$Date.Time)
```

```{r}
data$day <- factor(day(data$Date.Time))
data$month <- factor(month(data$Date.Time, label=TRUE))
data$year <- factor(year(data$Date.Time))
data$dayofweek <- factor(wday(data$Date.Time, label=TRUE))

data$hour <- factor(hour(hms(data$Time)))
data$minute <- factor(minute(hms(data$Time)))
data$second <- factor(second(hms(data$Time)))
```
Data Visualisation Part

Trips by hours in a day

```{r}
hourly_data <- data %>% group_by(hour) %>% dplyr::summarize(Total=n())

datatable(hourly_data)

ggplot(hourly_data, aes(hour, Total)) +
        geom_bar( stat = "identity", fill = "steelblue", color = "red") +
           ggtitle("Trips Every Hour") +
            theme(legend.position = "none") +
            scale_y_continuous(labels = comma)
```

Plotting Trips by hour and month
```{r}
month_hour <- data %>%
          group_by(month, hour) %>%
             dplyr::summarize(Total = n())

ggplot(month_hour, aes(hour, Total, fill = month)) + 
       geom_bar( stat = "identity") +
          ggtitle("Trips by Hour and Month") +
           scale_y_continuous(labels = comma)

```
Plotting trips by day of the month

```{r}
trips_by_day <- data %>% group_by(day) %>% dplyr::summarize(Trips=n())

ggplot(trips_by_day, aes(x= day, y= Trips))+ 
geom_bar(stat = "identity", fill = "steelblue") +
ggtitle("Trips by day of the month") + 
theme(legend.position = "none") + 
scale_y_continuous(labels = comma)

```
Now, plotting trips by day and month
```{r}

trips_month_days <- data %>% group_by(dayofweek, month) %>% dplyr::summarize(Trips=n())

ggplot(trips_month_days, aes(dayofweek, Trips, fill = month)) + 
geom_bar(stat = "identity", aes(fill = month), position = "dodge") + 
ggtitle("Trips by Day and Month") + 
scale_y_continuous(labels = comma) +
scale_fill_manual(values = c("red", "yellow", "black", "green", "grey", "brown"))

```
Trips by months

```{r}
monthly_data <-  data %>% group_by(month) %>% dplyr::summarize(Trips=n())

ggplot(monthly_data, aes(month, Trips, fill= month)) +
geom_bar(stat = "identity", aes(fill=month), position="dodge") +
ggtitle("Trips by months") +
scale_y_continuous(labels = comma) +
scale_fill_manual(values = c("red", "yellow", "black", "green", "grey", "brown"))
```
Heatmap Visualisation by hour and day.

```{r}
day_hour_data <- data %>% group_by(day, hour) %>% dplyr::summarize(Total = n())
datatable(day_hour_data)

ggplot(day_hour_data, aes(day, hour, fill = Total)) + 
geom_tile(color = "white") + 
ggtitle("Heat Map by Hour and Day")
```



